Irt plot extra

### 2022-02-22 manuscript.Rmd
Feedback round2: Typo updated in bodytext

A quantitative cross-sectional fully anonymous online survey was conducted via REDCap [@harris2009research; @harris2019redcap] by way of the University of Basel [sciCORE](https://redcap.scicore.unibas.ch). Participants were recruited via the [website](https://enlightenyourclock.org/participate-in-research) of a Comic co-released with the survey [@Weinzaepflen.2021], social media (i.e., LinkedIn, Twitter, Facebook), mailing lists, word of mouth, the investigators' personal contacts, and supported by distribution of the survey link via f.lux [@f.lux]. The landing page of the online survey had the explanatory statements where we mentioned participation was voluntary and that respondents could withdraw from participation anytime without being penalized. At the beginning of the survey, for the adult participants (>18 years) consent was recorded digitally. Underaged participants (<18 years) were urged to obtain assent from their parents/legal guardians. The survey took around 15 to 20 minutes and participants were not compensated. As a part of the demographic information participants provided information regarding age, sex, gender identity, occupational status, COVID-19 related occupational setting, time zone/country of residence and native language. The demographic characteristics of our sample are given in Table \@ref(tab:demotab).To ensure high data quality, five attention check items were included in the survey (e.g., "We want to make sure you are paying attention. What is 4+5?"). Participants needed to confirm that they were participating the survey for the first time. Questions incorporating retrospective recall were all aligned to the period of "past four weeks". At first we collected data for from `r nrow(EFA.descriptives)` participants (EFA sample). Next, we collected data from another `r nrow(CFA.descriptives)` participants (CFA sample) making a total sample of `r nrow(desctable.data)`. The data analysed in this study was collected between 17 May 2021 and 3 September 2021.

## Analytic Strategies

Figure \@ref(fig:FlowchartFig) summarizes the steps of our psychometric analysis. In our analysis we used R (version 4.1.0). In the item generation and selection phase we had developed a item pool of 48 items with six-point Likert type response format (0:Does not apply/I don't know; 1:Never, 2:Rarely; 3:Sometimes; 4:Often; 5:Always). Our purpose was to capture light exposure related behaviour. In that context, the first two response options: "Does not apply/I don't know" and "Never" were providing similar information. As such we collapsed them into one, making it a 5 point Likert type response format. We conducted an initial item analysis and proceed to the exploratory factor analysis (EFA) with all 48 items (EFA sample;n=`r nrow(EFA.descriptives)`). Prior to the EFA, necessary assumptions, including sample adequacy, normality assumptions, quality of correlation matrix were assessed. Our data violated both the univariate and multivariate normality assumptions. Due to these violations and the ordinal nature of our response data, in EFA we used  polychoric correlation matrix and employed principal axis (PA) as  factor extraction method with varimax rotation [@desjardinsHandbookEducationalMeasurement2018;@watkinsStepbyStepGuideExploratory2020]. We used a combination of factor identification method including Scree plot [@cattellScreeTestNumber1966], minimum average partials method [@velicerDeterminingNumberComponents1976], and Hull method [@lorenzo-sevaHullMethodSelecting2011] to identify factor numbers. To determine the latent structure, we followed the common guidelines : (i) no factors with fewer than three items (ii) no factors with a factor loading <0.3 (iii) no items with cross-loading > .3 across factors [@bandalosFactorAnalysisExploratory2018]. We also conducted a EFA on non-merged response options data (Supp. Table \@ref(tab:efaunmerged)) and rejected the latent structure obtained as the factors were less interpretable.

For reliability estimation we used internal consistency reliability coefficient ordinal alpha. Though  Cronbach's alpha coefficient is widely used for estimating internal consistency, it has a tendency to deflate the estimates for  Likert type data since the calculation is based on pearson-correlation matrix which requires response data to be continuous of nature [@gadermann2012estimating; @zumbo2007ordinal]. Subsequently to get  better estimates of reliability  we reported ordinal alpha for each factors using polychoric-correlation matrix [@zumbo2007ordinal]. We also estimated the internal consistency reliability of the total scale using McDonald's $\omega_t$  coefficient which is a better reliability estimate for multidimensional constructs [@dunn2014alpha; @sijtsma2009use]. Both ordinal alpha and McDonald's $\omega_t$ coefficient value range from 0 to 1 and higher value represents better reliability.

@@

We assessed the measurement invariance (MI) of our construct between native English speakers (n=`r invariance.data.descriptives.count[1,2]`) and non-non native English speakers (n=`r invariance.data.descriptives.count[2,2]`) among the CFA sample (n=`r nrow(CFA.descriptives)`). MI evaluates whether a construct has the psychometric equivalence and same meaning across groups [@putnick2016measurement;@klinePrinciplesPracticeStructural2015]. We used  structural equation modelling framework to assess the measurement invariance.  We successively compared four nested models:  configural, metric, scalar, and residual models using the $\chi^2$ difference test ($\Delta \chi^2$). Among MI models, configural model is the least restrictive model and residual model is the most restrictive model. A non-significant $\Delta \chi^2$ test between two nested measurement invariance models indicates mode fit does not significantly decrease for the superior model [@dimitrov2010testing] thus allowing the superior invariance model to be accepted. [@widaman1997exploring].

We also analysed possible semantic overlap of our developed tool using "Semantic Scale Network" (SSN) engine [@rosenbusch2020semantic]. The SSN detects semantically related scales and provides cosine similarity index ranging between -.66 to 1 [@rosenbusch2020semantic]. Pair of scales with a cosine similarity index value of 1 indicates they are perfectly semantically similar scales indicating redundancy. Additionally, To identify the educational grade level required to understand the items in our tool we subjected the tool to Flesch-Kincaid Grade Level [@flesch1948new]

@@

"Semantic Scale Network"(SSN) analysis [@rosenbusch2020semantic] indicated that LEBA (23 items)  appeared most strongly related to scales about sleep: "Sleep Disturbance Scale For Children" [@bruni1996sleep] and "WHO-Composite International Diagnostic Interview (CIDI): Insomnia"[@who1990].The cosine similarities lie between .47 to .51. Flesch-Kincaid Grade Level [@flesch1948new] analysis on the the 23 items of our scale indicated required educational grade level was 3.33 and with a age above 8.33.

@@

Table \@ref(tab:itemfittab) summarizes the item fit indexes of the items. All of the items had RMSEA value $\le$.06 indicating adequate fit of the items. Sup.Figure \@ref(fig:personfit) depicts the person fit Zh statistics histogram of out fitted models. Zh statistics are larger than -2 for most participants, suggesting a good person fit of the selected IRT models.

@@

A CFA on a separate sample ((n=`r nrow(CFA.descriptives)` gave a five-factor solution (CFI =`r apa(fit.2[5],2,F)`; TLI = `r apa(fit.2[6],2,F)`); RMSEA = `r apa(fit.2[7],2,F)`[`r apa(fit.2[8],2,F)`-`r apa(fit.2[9],2,F)`, 90% CI]) and SRMR = `r apa(fit.2[10],2,F)`) after discarding two item. The internal consistency McDonald's $\omega_t$ of the five factors were  satisfactory (`r apa(rel.k[2,1],2,F)`, `r apa(rel.k[2,2],2,F)`, `r apa(rel.k[2,3],2,F)`, `r apa(rel.k[2,4],2,F)`, `r apa(rel.k[2,5],2,F)`) Internal consistency reliability of the total scale (23 items) was also satisfactory, McDonald's $\omega_t$= `r apa(rel.k[5,6],2,F)`.In the same sample, our measurement invariance analysis  revealed that the latent structure attained the residual measurement invariance across subgroups: male and female (CFI: `r apa (residual.fitm[5,1],2,F)`, TLI: `r apa (residual.fitm[6,1],2,F)`, SRMR: `r apa(residual.fitm[6,1],2,F)`).

The "Semantic Scale Network"(SSN) analysis [@rosenbusch2020semantic] on the retained 23 items showed "LEBA" was  related to  "Sleep Disturbance Scale For Children" (SDSC) [@bruni1996sleep] and "WHO-Composite International Diagnostic Interview (CIDI): Insomnia"[@who1990]. Upon inspecting the item contents we found items under "Using phone and smart-watch in bed" and "Using light before bedtime" have semantic overlap with the items of SDSC ans CIDI. Items in those two scales were looking into behaviours related to sleep. As such the similarity index obtained is expected. Flesch-Kincaid Grade Level [@flesch1948new] analysis on the the 23 items of our scale indicated required educational grade level was 3.33 and with a age above 8.33.

We developed a short-LEBA (18 items) using IRT analysis. We fitted a graded response model model to the combined EFA and CFA sample (n=`r nrow(data)`). We discarded five items with relatively flat item information curve [I($\theta$) <.20]. IRT analysis indicated short form of LEBA is a psychometrically sound measure. Item fit indexes and person fit index for all five fitted model were acceptable. Items had diverse slope parameters indicating a good range of discrimination- the ability to differentiate respondents with different levels of the light exposure related behaviour. Test information curve also indicated a good coverage of underlying trait continuum with precision.

LEBA can be used to profile individuals based on their light exposure related behaviours, which can facilitate the development process of individual interventions to promote health. All the five factors of LEBA can identify 'problematic' behaviours that are opposed to good light hygiene.

@@

We developed a novel self-reported subjective tool- "Light exposure behaviour assessment"(LEBA) to capture light exposure related behaviour. We developed 48 items, judged the relevance and content of the items and conducted a large scale geographically unrestricted cross-sectional survey. Our EFA gave a five solution with 25 items. A CFA with this 25-item scale again offered a five-factor solution, but this time two more item was discarded. The 23-item "LEBA" was found reliable (internal consistency) and valid (structural validity). A short-form of LEBA was developed using IRT analysis. IRT analysis gave a 18-item scale with  a good  coverage across the underlying trait continuum. Hence, we could recommend that LEBA can be used to measure different aspects of light exposure related behaviour.

>>>>

A quantitative cross-sectional fully anonymous online survey was conducted via REDCap [@harris2009research; @harris2019redcap] by way of the University of Basel [sciCORE](https://redcap.scicore.unibas.ch). Participants were recruited via the [website](https://enlightenyourclock.org/participate-in-research) (https://enlightenyourclock.org/participate-in-research) of the science-communication comic-book "Enlighten your clock" co-released with the survey [@Weinzaepflen.2021], social media (i.e., LinkedIn, Twitter, Facebook), mailing lists, word of mouth, the investigators' personal contacts, and supported by distribution of the survey link via f.lux [@f.lux]. The landing page of the online survey had the explanatory statements where we mentioned participation was voluntary and that respondents could withdraw from participation anytime without being penalized. At the beginning of the survey, for the adult participants (>18 years) consent was recorded digitally. Underaged participants (<18 years) were urged to obtain assent from their parents/legal guardians. The entire survey was estimated to take <30 minutes. Participants were not compensated. As a part of the demographic information participants provided information regarding age, sex, gender identity, occupational status, COVID-19 related occupational setting, time zone/country of residence and native language. The demographic characteristics of our sample are given in Table \@ref(tab:demotab).To ensure high data quality, five attention check items were included in the survey (e.g., "We want to make sure you are paying attention. What is 4+5?"). Participants were asked to confirm that they were participating the survey for the first time. Questions incorporating retrospective recall were all aligned to the period of "past four weeks". At first we collected data for from `r nrow(EFA.descriptives)` participants (EFA sample). Next, we collected data from another `r nrow(CFA.descriptives)` participants (CFA sample) making a total sample of `r nrow(desctable.data)`. The data analysed in this study was collected between 17 May 2021 and 3 September 2021.

## Analytic Strategy

Figure \@ref(fig:FlowchartFig) summarizes the steps of our psychometric analysis. In our analysis we used R (version 4.1.0) [@R-base]. In the item generation and selection phase we had developed a item pool of 48 items with six-point Likert type response format (0-Does not apply/I don't know, 1-Never, 2-Rarely 3-Sometimes, 4-Often, 5-Always). Our purpose was to capture light exposure related behaviour. In that context, the first two response options: "Does not apply/I don't know" and "Never" were providing similar information. As such we collapsed them into one, making it a 5 point Likert type response format. We conducted an initial item analysis and proceed to the exploratory factor analysis (EFA) with all 48 items (EFA sample;n=`r nrow(EFA.descriptives)`). Prior to the EFA, necessary assumptions, including sample adequacy, normality assumptions, quality of correlation matrix were assessed. Our data violated both the univariate and multivariate normality assumptions. Due to these violations and the ordinal nature of our response data, in EFA we used  polychoric correlation matrix and employed principal axis (PA) as  factor extraction method with varimax rotation [@desjardinsHandbookEducationalMeasurement2018;@watkinsStepbyStepGuideExploratory2020]. We used a combination of factor identification method including Scree plot [@cattellScreeTestNumber1966], minimum average partials method [@velicerDeterminingNumberComponents1976], and Hull method [@lorenzo-sevaHullMethodSelecting2011] to identify factor numbers. To determine the latent structure, we followed the common guidelines : (i) no factors with fewer than three items (ii) no factors with a factor loading <0.3 (iii) no items with cross-loading > .3 across factors [@bandalosFactorAnalysisExploratory2018]. We also conducted a EFA on non-merged response options data (Supp. Table \@ref(tab:efaunmerged)).

For reliability estimation we used internal consistency reliability coefficient ordinal $\alpha$. Though Cronbach's alpha coefficient is widely used for estimating internal consistency, it has a tendency to deflate the estimates for Likert-type data since the calculation is based on pearson-correlation matrix which requires response data to be continuous of nature [@gadermann2012estimating; @zumbo2007ordinal]. Subsequently to get  better estimates of reliability  we reported ordinal alpha for each factors using polychoric-correlation matrix [@zumbo2007ordinal]. We also estimated the internal consistency reliability of the total scale using McDonald's $\omega_t$  coefficient which is a better reliability estimate for multidimensional constructs [@dunn2014alpha; @sijtsma2009use]. Both ordinal alpha and McDonald's $\omega_t$ coefficient value range from 0 to 1 and higher value represents better reliability.

@@

We assessed the measurement invariance (MI) of our construct between native English speakers (n=`r invariance.data.descriptives.count[1,2]`) and non native English speakers (n=`r invariance.data.descriptives.count[2,2]`) among the CFA sample (n=`r nrow(CFA.descriptives)`). MI evaluates whether a construct has the psychometric equivalence and same meaning across groups [@putnick2016measurement;@klinePrinciplesPracticeStructural2015]. We used  structural equation modelling framework to assess the measurement invariance.  We successively compared four nested models:  configural, metric, scalar, and residual models using the $\chi^2$ difference test ($\Delta \chi^2$). Among MI models, configural model is the least restrictive model and residual model is the most restrictive model. A non-significant $\Delta \chi^2$ test between two nested measurement invariance models indicates mode fit does not significantly decrease for the superior model [@dimitrov2010testing] thus allowing the superior invariance model to be accepted. [@widaman1997exploring].

We also analysed possible semantic overlap of our developed tool using "Semantic Scale Network" (SSN) engine [@rosenbusch2020semantic]. The SSN detects semantically related scales and provides cosine similarity index ranging between -.66 to 1 [@rosenbusch2020semantic]. Pair of scales with a cosine similarity index value of 1 indicates they are perfectly semantically similar scales indicating redundancy. Additionally, to identify the educational grade level required to understand the items in our tool we subjected the tool to Flesch-Kincaid Grade Level [@flesch1948new]

@@

"Semantic Scale Network"(SSN) analysis [@rosenbusch2020semantic] indicated that LEBA (23 items)  appeared most strongly related to scales about sleep: "Sleep Disturbance Scale For Children" [@bruni1996sleep] and "WHO-Composite International Diagnostic Interview (CIDI): Insomnia"[@who1990]. The cosine similarities lie between .47 to .51. Flesch-Kincaid Grade Level [@flesch1948new] analysis on the the 23 items of our scale indicated required educational grade level was 3.33 and with a age above 8.33.

@@

Table \@ref(tab:itemfittab) summarizes the item fit indexes of the items. All of the items had RMSEA value $\le$.06 indicating adequate fit of the items to the fitted IRT model. Sup.Figure \@ref(fig:personfit) depicts the person fit Zh statistics histogram of out fitted models. Zh statistics are larger than -2 for most participants, suggesting a good person fit of the selected IRT models.

@@

<!-- A CFA on a separate sample ((n=`r nrow(CFA.descriptives)` gave a five-factor solution (CFI =`r apa(fit.2[5],2,F)`; TLI = `r apa(fit.2[6],2,F)`); RMSEA = `r apa(fit.2[7],2,F)`[`r apa(fit.2[8],2,F)`-`r apa(fit.2[9],2,F)`, 90% CI]) and SRMR = `r apa(fit.2[10],2,F)`) after discarding two item. The internal consistency McDonald's $\omega_t$ of the five factors were  satisfactory (`r apa(rel.k[2,1],2,F)`, `r apa(rel.k[2,2],2,F)`, `r apa(rel.k[2,3],2,F)`, `r apa(rel.k[2,4],2,F)`, `r apa(rel.k[2,5],2,F)`) Internal consistency reliability of the total scale (23 items) was also satisfactory, McDonald's $\omega_t$= `r apa(rel.k[5,6],2,F)`.In the same sample, our measurement invariance analysis  revealed that the latent structure attained the residual measurement invariance across subgroups: male and female (CFI: `r apa (residual.fitm[5,1],2,F)`, TLI: `r apa (residual.fitm[6,1],2,F)`, SRMR: `r apa(residual.fitm[6,1],2,F)`). -->

<!-- The "Semantic Scale Network"(SSN) analysis [@rosenbusch2020semantic] on the retained 23 items showed "LEBA" was  related to  "Sleep Disturbance Scale For Children" (SDSC) [@bruni1996sleep] and "WHO-Composite International Diagnostic Interview (CIDI): Insomnia"[@who1990]. Upon inspecting the item contents we found items under "Using phone and smart-watch in bed" and "Using light before bedtime" have semantic overlap with the items of SDSC ans CIDI. Items in those two scales were looking into behaviours related to sleep. As such the similarity index obtained is expected. Flesch-Kincaid Grade Level [@flesch1948new] analysis on the the 23 items of our scale indicated required educational grade level was 3.33 and with a age above 8.33. -->

<!-- We developed a short-LEBA (18 items) using IRT analysis. We fitted a graded response model model to the combined EFA and CFA sample (n=`r nrow(data)`). We discarded five items with relatively flat item information curve [I($\theta$) <.20]. IRT analysis indicated short form of LEBA is a psychometrically sound measure. Item fit indexes and person fit index for all five fitted model were acceptable. Items had diverse slope parameters indicating a good range of discrimination- the ability to differentiate respondents with different levels of the light exposure related behaviour. Test information curve also indicated a good coverage of underlying trait continuum with precision. -->

LEBA can be used to profile individuals based on their light exposure related behaviours, which can facilitate the development process of individual interventions to promote health. All the five factors of LEBA may identify 'problematic' behaviours that are opposed to good light hygiene.

@@

We developed a novel self-reported subjective tool-"Light exposure behaviour assessment"(LEBA) to capture light exposure related behaviour. We developed 48 items, judged the relevance and content of the items and conducted a large scale geographically unrestricted cross-sectional survey. Our EFA gave a five solution with 25 items. A CFA with this 25-item scale again offered a five-factor solution, but this time two more item was discarded. The 23-item "LEBA" was found reliable (internal consistency) and valid (structural validity). A short-form of LEBA was developed using IRT analysis. IRT analysis gave a 18-item scale with  a good  coverage across the underlying trait continuum. Hence, we could recommend that LEBA can be used to measure different aspects of light exposure related behaviour.
